Decoding Whale Songs Using Backtranslation and Transformer-Based Models

### 2024-08-04 tex/paper.tex
fixed Ket not existing in dictionary and failing spellcheck by adding to dictionary.txt

\begin{itemize}
    \item $P_{we}$: English likelihood penalty
    \item $P_{wew}$: Whale context probability penalty
    \item $P_{sd}$: Semantic distance penalty
\end{itemize}
\subsection{English Likelihood Penalty (\texorpdfstring{\(P_{we}\)}{Pwe})}
$P_{we}$ is calculated using $M_e$ to evaluate the average probability per token of the \textit{whale\_to\_english} paragraph. A high $P_{we}$ indicates an unlikely English paragraph.

\subsection{Whale Context Probability Penalty (\texorpdfstring{\(P_{wew}\)}{Pwew})}

$P_{wew}$ is determined by inserting \textit{whale\_to\_english\_back\_to\_whale} into the overall whale text and calculating the probability using $M_w$. A high $P_{wew}$ indicates low probability of the translated whale text appearing in that context.

\subsection{Semantic Distance Penalty (\texorpdfstring{\(P_{sd}\)}{Psd})}
$P_{sd}$ measures the semantic distance between the original whale paragraph and \textit{whale\_to\_english\_back\_to\_whale}. A high $P_{sd}$ indicates significant divergence in meaning. This penalty is weighted higher than $P_{wew}$ to prioritize meaning preservation.

@@

"In addition, our results show the
feasibility of using "self-supervised" learning on proxy tasks and applying trained neural networks to label unseen
coda data according to type, clan, and individual whale, which could drastically expedite the analysis of recorded
sperm whale signals.
The datasets in this study collectively contain ~26,000 annotated sperm whale codas and represent the largest
available labeled data of sperm whale coda signals. Using these datasets, this study exhibits the potential appli-
cations of ML to sperm whale vocalizations by demonstrating the ability of NNs to perform detection and clas-
sification tasks with high degrees of accuracy"
This dataset is what is run to classify, althoug:
https://github.com/dgruber212/Sperm_Whale_Machine_Learning/tree/master

@@

These datasets allow testing when converting from a given spoken language to a series of letters corresponding to vocals, similar to Navajo. It helps establish a baseline for the ability of an unsupervised translation algorithm to effectively translate an unknown language into a known language.
\section{Potential Challenges}
\begin{enumerate}
    \item Poor performance of $M_w$
    \item Insufficient context for accurate translation
    \item Tendency towards common whale utterances in back-translation
    \item Development of a system to "cheat" the translation process
    \item Unclear direction of penalties for efficient learning
    \item Overfitting to specific training data
    \item Risk of translating only sounds and grammar without capturing meaning
\end{enumerate}
\section{Miscellaneous end notes}

>>>>

\begin{itemize}
    \item $P_{we}$: English likelihood penalty
    \item $P_{we}$: Whale context probability penalty
    \item $P_{sd}$: Semantic distance penalty
\end{itemize}
\subsection{English Likelihood Penalty (\texorpdfstring{\(P_{we}\)}{Pwe})}
$P_{we}$ is calculated using $M_e$ to evaluate the average probability per token of the \textit{whale\_to\_english} paragraph. A high $P_{we}$ indicates an unlikely English paragraph.

\subsection{Whale Context Probability Penalty (\texorpdfstring{\(P_{we}\)}{Pwew})}

$P_{we}$ is determined by inserting \textit{whale\_to\_english\_back\_to\_whale} into the overall whale text and calculating the probability using $M_w$. A high $P_{we}$ indicates low probability of the translated whale text appearing in that context.

\subsection{Semantic Distance Penalty (\texorpdfstring{\(P_{sd}\)}{Psd})}
$P_{sd}$ measures the semantic distance between the original whale paragraph and \textit{whale\_to\_english\_back\_to\_whale}. A high $P_{sd}$ indicates significant divergence in meaning. This penalty is weighted higher than $P_{we}$ to prioritize meaning preservation.

@@

"In addition, our results show the
feasibility of using "self-supervised" learning on proxy tasks and applying trained neural networks to label unseen
coda data according to type, clan, and individual whale, which could drastically expedite the analysis of recorded
sperm whale signals.
The datasets in this study collectively contain ~26,000 annotated sperm whale codas and represent the largest
available labeled data of sperm whale coda signals. Using these datasets, this study exhibits the potential appli-
cations of ML to sperm whale vocalizations by demonstrating the ability of NNs to perform detection and clas-
sification tasks with high degrees of accuracy"
This dataset is what is run to classify, although:
https://github.com/dgruber212/Sperm_Whale_Machine_Learning/tree/master

@@

These datasets allow testing when converting from a given spoken language to a series of letters corresponding to vocals, similar to Navajo. It helps establish a baseline for the ability of an unsupervised translation algorithm to effectively translate an unknown language into a known language.
\section{Potential Challenges}
\begin{enumerate}
    \item Poor performance of $M_w$.
    \item Insufficient context for accurate translation.
    \item Tendency towards common whale utterances in back-translation.
    \item Development of a system to "cheat" the translation process.
    \item Unclear direction of penalties for efficient learning.
    \item Overfitting to specific training data.
    \item Risk of translating only sounds and grammar without capturing meaning.
\end{enumerate}
\section{Miscellaneous end notes}
